parental migration and adolescents educational performance

### 2022-06-12 data/Gu-thesis-script.Rmd
more dependent var: raw score, z-score, std score, difference to mean

```{r}
d <- d %>% mutate(
  age = 2013 - a02a, #student's age at baseline
  local.resi = age - a07, #years of residence at the current county
  w1chn_100 = tr_chn/w2upchn*100, #raw exam scores at 100-point scale
  w1mat_100 = tr_mat/w2upmat*100,
  w1eng_100 = tr_eng/w2upeng*100,
  w2chn_100 = w2chn/w2upchn*100,
  w2mat_100 = w2mat/w2upmat*100,
  w2eng_100 = w2eng/w2upeng*100
)
d <- d %>% mutate(
  w1ttl_100 = w1chn_100 + w1mat_100 + w1eng_100,
  w2ttl_100 = w2chn_100 + w2mat_100 + w2eng_100,
  )
```

Detect outliers
```{r}
d_outlier <- d %>% filter(w1chn_100 > 100 | w1mat_100>100|w1eng_100>100|w2chn_100>100|w2mat_100>100|w2eng_100>100)
d1 <- d %>% anti_join(d_outlier, by=c('ids', 'clsids', 'schids', 'ctyids'))
```

@@

Sample
```{r}
d_cl <- d1 %>% 
  filter(w2a0801==1) %>% #parents who are married (and alive)
  filter(be1201==1 & be1202==1) %>% #both parents live with children at baseline
  filter(stmigrant==1) #having stayed in the current county for at least six years (duration of primary school) before baseline
```

## Variables 
```{r}
d_cl <- d_cl %>% 
  mutate(chn_dif = w2chn_100 - w1chn_100,
         mat_dif = w2mat_100 - w1mat_100,
         eng_dif = w2eng_100 - w1eng_100,
         ttl_dif = w2ttl_100 - w1ttl_100,
         cog_dif = w2cog3pl - cog3pl) %>%
  drop_na(cog_dif, chn_dif, mat_dif, eng_dif, ttl_dif)
```

Control variables
```{r}
d_cl <- d_cl %>% mutate(
         ch.girl = if_else(a01 == 2, 1, 0), #student is girl
         ch.rural = sthktype, #student holds rural (agricultural) hukou
         #ch.only.child = if_else(b01 == 1, 1, 0), #student is only child
         ch.boarding = stbrd, #student is boarding at school
         ch.ethn.minority = if_else(a03 != 1, 1, 0), #student is ethnic minority
         ch.health = a17, #student's self-reported health
         ch.skip.grd = if_else(c03 == 0, 0, 1),
         ch.repeat.grd = if_else(c04 == 0, 0, 1),
         ch.illness = if_else(a10 == 1, 1, 0),
         ch.preschool = if_else(c01 == 1, 1, 0),
         par.edu = stprhedu, #Among two parents, the best-educated one's educational level
         par.ccp = if_else(be06 == 1, 1, 0), #At least one parent is communist party member
         home.finances = steco_3c,
         home.internet = if_else(b13 == 2, 1, 0), #student has internet access at home
         home.book = b12 #Amount of book at home, student self-rated 
) %>% drop_na(
ch.girl, ch.rural, ch.health, ch.boarding, ch.ethn.minority, ch.health, ch.skip.grd, ch.repeat.grd, ch.illness, ch.preschool, par.edu, par.ccp, home.internet, home.book, home.finances)
```

@@

```{r}
did_unres_ttl <- feols(ttl_dif ~ migr + w1ttl_100 | schids, data = d_cl) 
did_unres_ttl_m_only <- feols(ttl_dif ~ migr_mo_only + w1ttl_100 | schids, data = d_cl) 
did_unres_ttl_m <- feols(ttl_dif ~ migr_mo + w1ttl_100 | schids, data = d_cl) 
did_unres_ttl_f_only <- feols(ttl_dif ~ migr_fa_only + w1ttl_100 | schids, data = d_cl) 
did_unres_ttl_f <- feols(ttl_dif ~ migr_fa + w1ttl_100 | schids, data = d_cl) 
did_unres_ttl_both <- feols(ttl_dif ~ migr_both + w1ttl_100 | schids, data = d_cl)
```

@@

Regressing on Chinese exam score
```{r}
did_unres_chn <- feols(chn_dif ~ migr + w1chn_100 | schids, data = d_cl) 
did_unres_chn_m_only <- feols(chn_dif ~ migr_mo_only + w1chn_100 | schids, data = d_cl) 
did_unres_chn_m <- feols(chn_dif ~ migr_mo + w1chn_100 | schids, data = d_cl) 
did_unres_chn_f_only <- feols(chn_dif ~ migr_fa_only + w1chn_100 | schids, data = d_cl) 
did_unres_chn_f <- feols(chn_dif ~ migr_fa + w1chn_100 | schids, data = d_cl) 
did_unres_chn_both <- feols(chn_dif ~ migr_both + w1chn_100 | schids, data = d_cl)
```

@@

Regressing on Chinese exam score
```{r}
did_adj_unres_chn <- feols(chn_dif ~ migr + w1chn_100  + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_chn_m_only <- feols(chn_dif ~ migr_mo_only + w1chn_100  + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_chn_m <- feols(chn_dif ~ migr_mo + w1chn_100  + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_chn_f_only <- feols(chn_dif ~ migr_fa_only + w1chn_100  + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_chn_f <- feols(chn_dif ~ migr_fa + w1chn_100  + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_chn_both <- feols(chn_dif ~ migr_both + w1chn_100  + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl)
```

>>>>

```{r}
d <- d %>% mutate(
  age = 2013 - a02a, #student's age at baseline
  local.resi = age - a07, #years of residence at the current county
  w1chn_100 = tr_chn/w2upchn*100, #raw exam scores at 100-point scale
  w1mat_100 = tr_mat/w2upmat*100,
  w1eng_100 = tr_eng/w2upeng*100,
  w2chn_100 = w2chn/w2upchn*100,
  w2mat_100 = w2mat/w2upmat*100,
  w2eng_100 = w2eng/w2upeng*100
)
d <- d %>% mutate(
  w1ttl = tr_chn + tr_mat + tr_eng,
  w2ttl = w2chn + w2mat + w2eng,
  w1ttl_100 = w1chn_100 + w1mat_100 + w1eng_100,
  w2ttl_100 = w2chn_100 + w2mat_100 + w2eng_100,
  )
```

```{r}
d1 <- d %>% filter(w2ttl_100<=300 & w1ttl_100<=300) #remove outliers with total scores greater than 300
```

within-class z score
```{r}
d1 <- d1 %>% group_by(clsids) %>%
  mutate(
    w1ttl_z = as.numeric(scale(tr_chn + tr_mat + tr_eng)),
    w2ttl_z = as.numeric(scale(w2chn + w2mat + w2eng))
  ) %>% ungroup()
```

```{r}
d1 <- d1 %>% group_by(clsids) %>%
  mutate(
    w1ttl_max = max(w1ttl, na.rm = T),
    w1ttl_min = min(w1ttl, na.rm = T),
    w2ttl_max = max(w2ttl, na.rm = T),
    w2ttl_min = min(w2ttl, na.rm = T),
    w1ttl_100_mean = mean(w1ttl_100, na.rm = T),
    w2ttl_100_mean = mean(w2ttl_100, na.rm = T)
  ) %>% ungroup()
d1 <- d1 %>%
  mutate(
    w1ttl_std = (w1ttl - w1ttl_min)/(w1ttl_max - w1ttl_min),
    w2ttl_std = (w2ttl - w2ttl_min)/(w2ttl_max - w2ttl_min),
    w1ttl_100_dm = (w1ttl_100 - w1ttl_100_mean),
    w2ttl_100_dm = (w2ttl_100 - w2ttl_100_mean)
  )
```

Detect outliers
```{}
d_outlier <- d %>% filter(w1chn_100 > 100 | w1mat_100>100|w1eng_100>100|w2chn_100>100|w2mat_100>100|w2eng_100>100)
d_outlier1 <- d_outlier %>% select(clsids, schids, tr_chn, tr_mat, tr_eng, w2chn, w2upchn, w2mat, w2upmat, w2eng, w2upeng)
d1 <- d %>% anti_join(d_outlier, by=c('ids', 'clsids', 'schids', 'ctyids'))
```

@@

```{r}
d2 <- d1 %>% 
  mutate(chn_dif = w2chn_100 - w1chn_100,
         mat_dif = w2mat_100 - w1mat_100,
         eng_dif = w2eng_100 - w1eng_100,
         ttl_dif = w2ttl_100 - w1ttl_100,
         ttl_z_dif = w2ttl_z - w1ttl_z,
         ttl_std_dif = w2ttl_std - w1ttl_std,
         ttl_dm_dif = w2ttl_100_dm - w1ttl_100_dm,
         cog_dif = w2cog3pl - cog3pl) %>%
  drop_na(cog_dif)
```

Sample
```{r}
d_cl <- d2 %>% 
  filter(w2a0801==1) %>% #keep parents who are married (and alive)
  filter(be1201==1 & be1202==1) %>% #keep both parents live with children at baseline
  filter(stmigrant==1) #keep local children 
```

## Variables

Control variables
```{r}
d_cl <- d_cl %>% mutate(
         ch.girl = if_else(a01 == 2, 1, 0), #student is girl
         ch.rural = sthktype, #student holds rural (agricultural) hukou
         #ch.only.child = if_else(b01 == 1, 1, 0), #student is only child
         ch.boarding = stbrd, #student is boarding at school
         ch.ethn.minority = if_else(a03 != 1, 1, 0), #student is ethnic minority
         ch.health = a17, #student's self-rated health
         ch.skip.grd = if_else(c03 == 0, 0, 1), #student has skipped grade
         ch.repeat.grd = if_else(c04 == 0, 0, 1), #student has repeated grade
         ch.illness = if_else(a10 == 1, 1, 0), #student had serious illness before primary school
         ch.preschool = if_else(c01 == 1, 1, 0), #student had attended preschool
         par.edu = stprhedu, #Among two parents, the best-educated one's educational level
         par.ccp = if_else(be06 == 1, 1, 0), #At least one parent is communist party member
         home.finances = steco_5c, #self-rated household financial situation
         home.internet = if_else(b13 == 2, 1, 0), #student has internet access at home
         home.book = b12 #Amount of book at home, student self-rated 
) %>% drop_na(
ch.girl, ch.rural, ch.health, ch.boarding, ch.ethn.minority, ch.health, ch.skip.grd, ch.repeat.grd, ch.illness, ch.preschool, par.edu, par.ccp, home.internet, home.book, home.finances)
```

@@

raw score
```{r}
did_unres_ttl <- feols(ttl_dif ~ migr + w1ttl_100 | schids, data = d_cl) 
did_unres_ttl_m_only <- feols(ttl_dif ~ migr_mo_only + w1ttl_100 | schids, data = d_cl) 
did_unres_ttl_m <- feols(ttl_dif ~ migr_mo + w1ttl_100 | schids, data = d_cl) 
did_unres_ttl_f_only <- feols(ttl_dif ~ migr_fa_only + w1ttl_100 | schids, data = d_cl) 
did_unres_ttl_f <- feols(ttl_dif ~ migr_fa + w1ttl_100 | schids, data = d_cl) 
did_unres_ttl_both <- feols(ttl_dif ~ migr_both + w1ttl_100 | schids, data = d_cl)
```

@@

z score
```{r}
did_unres_ttl_z <- feols(ttl_z_dif ~ migr + w1ttl_z | schids, data = d_cl) 
did_unres_ttl_z_m_only <- feols(ttl_z_dif ~ migr_mo_only + w1ttl_z | schids, data = d_cl) 
did_unres_ttl_z_m <- feols(ttl_z_dif ~ migr_mo + w1ttl_z | schids, data = d_cl) 
did_unres_ttl_z_f_only <- feols(ttl_z_dif ~ migr_fa_only + w1ttl_z | schids, data = d_cl) 
did_unres_ttl_z_f <- feols(ttl_z_dif ~ migr_fa + w1ttl_z | schids, data = d_cl) 
did_unres_ttl_z_both <- feols(ttl_z_dif ~ migr_both + w1ttl_z | schids, data = d_cl)
```

```{r}
models_unres_ttl_z <- list(
'Any Parent Absent' = did_unres_ttl_z,
'Only Mother Absent' = did_unres_ttl_z_m_only,
'Mother Absent' = did_unres_ttl_z_m,
'Only Father Absent' = did_unres_ttl_z_f_only,
'Father Absent' = did_unres_ttl_z_f,
'Both Parents Absent' = did_unres_ttl_z_both) 
table_unres_ttl_z <- models_unres_ttl_z %>% msummary(stars = c('*' = .05, '**' = .01, '***' = .001), 
                              fmt = 2,
                              coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "w1ttl_dif" = "Baseline difference"), output = 'markdown')
table_unres_ttl_z
```

std
```{r}
did_unres_ttl_std <- feols(ttl_std_dif ~ migr + w1ttl_std | schids, data = d_cl) 
did_unres_ttl_std_m_only <- feols(ttl_std_dif ~ migr_mo_only + w1ttl_std | schids, data = d_cl) 
did_unres_ttl_std_m <- feols(ttl_std_dif ~ migr_mo + w1ttl_std | schids, data = d_cl) 
did_unres_ttl_std_f_only <- feols(ttl_std_dif ~ migr_fa_only + w1ttl_std | schids, data = d_cl) 
did_unres_ttl_std_f <- feols(ttl_std_dif ~ migr_fa + w1ttl_std | schids, data = d_cl) 
did_unres_ttl_std_both <- feols(ttl_std_dif ~ migr_both + w1ttl_std | schids, data = d_cl)
```

```{r}
models_unres_ttl_std <- list(
'Any Parent Absent' = did_unres_ttl_std,
'Only Mother Absent' = did_unres_ttl_std_m_only,
'Mother Absent' = did_unres_ttl_std_m,
'Only Father Absent' = did_unres_ttl_std_f_only,
'Father Absent' = did_unres_ttl_std_f,
'Both Parents Absent' = did_unres_ttl_std_both) 
table_unres_ttl_std <- models_unres_ttl_std %>% msummary(stars = c('*' = .05, '**' = .01, '***' = .001), 
                              fmt = 2,
                              coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "w1ttl_dif" = "Baseline difference"), output = 'markdown')
table_unres_ttl_std
```

dif to mean
```{r}
did_unres_ttl_dm <- feols(ttl_dm_dif ~ migr + w1ttl_100_dm | schids, data = d_cl) 
did_unres_ttl_dm_m_only <- feols(ttl_dm_dif ~ migr_mo_only + w1ttl_100_dm | schids, data = d_cl) 
did_unres_ttl_dm_m <- feols(ttl_dm_dif ~ migr_mo + w1ttl_100_dm | schids, data = d_cl) 
did_unres_ttl_dm_f_only <- feols(ttl_dm_dif ~ migr_fa_only + w1ttl_100_dm | schids, data = d_cl) 
did_unres_ttl_dm_f <- feols(ttl_dm_dif ~ migr_fa + w1ttl_100_dm | schids, data = d_cl) 
did_unres_ttl_dm_both <- feols(ttl_dm_dif ~ migr_both + w1ttl_100_dm | schids, data = d_cl)
```

```{r}
models_unres_ttl_dm <- list(
'Any Parent Absent' = did_unres_ttl_dm,
'Only Mother Absent' = did_unres_ttl_dm_m_only,
'Mother Absent' = did_unres_ttl_dm_m,
'Only Father Absent' = did_unres_ttl_dm_f_only,
'Father Absent' = did_unres_ttl_dm_f,
'Both Parents Absent' = did_unres_ttl_dm_both) 
table_unres_ttl_dm <- models_unres_ttl_dm %>% msummary(stars = c('*' = .05, '**' = .01, '***' = .001), 
                              fmt = 2,
                              coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "w1ttl_dif" = "Baseline difference"), output = 'markdown')
table_unres_ttl_dm
```

### DID by subject
Regressing on Chinese exam score
```{r}
did_unres_chn <- feols(chn_dif ~ migr + w1chn_100 | schids, data = d_cl) 
did_unres_chn_m_only <- feols(chn_dif ~ migr_mo_only + w1chn_100 | schids, data = d_cl) 
did_unres_chn_m <- feols(chn_dif ~ migr_mo + w1chn_100 | schids, data = d_cl) 
did_unres_chn_f_only <- feols(chn_dif ~ migr_fa_only + w1chn_100 | schids, data = d_cl) 
did_unres_chn_f <- feols(chn_dif ~ migr_fa + w1chn_100 | schids, data = d_cl) 
did_unres_chn_both <- feols(chn_dif ~ migr_both + w1chn_100 | schids, data = d_cl)
```

@@

Regressing on total exam score
```{r}
did_adj_unres_ttl_z <- feols(ttl_z_dif ~ migr + w1ttl_z + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_ttl_z_m_only <- feols(ttl_z_dif ~ migr_mo_only + w1ttl_z + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_ttl_z_m <- feols(ttl_z_dif ~ migr_mo + w1ttl_z + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_ttl_z_f <- feols(ttl_z_dif ~ migr_fa + w1ttl_z + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_ttl_z_f_only <- feols(ttl_z_dif ~ migr_fa_only + w1ttl_z + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_ttl_z_both <- feols(ttl_z_dif ~ migr_both + w1ttl_z + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl)
```

Results
```{r}
models_adj_unres_ttl_z <- list(
'Any Parent Absent' = did_adj_unres_ttl_z,
'Only Mother Absent' = did_adj_unres_ttl_z_m_only,
'Mother Absent' = did_adj_unres_ttl_z_m,
'Only Father Absent' = did_adj_unres_ttl_z_f_only,
'Father Absent' = did_adj_unres_ttl_z_f,
'Both Parents Absent' = did_adj_unres_ttl_z_both) 
table_adj_unres_ttl_z <- models_adj_unres_ttl_z %>% msummary(stars = c('*' = .05, '**' = .01, '***' = .001), fmt = 2, coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "w1ttl_dif" = "Baseline difference"), output = 'markdown')
table_adj_unres_ttl_z
```

### DID by subject
Regressing on Chinese exam score
```{r}
did_adj_unres_chn <- feols(chn_dif ~ migr + w1chn_100  + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_chn_m_only <- feols(chn_dif ~ migr_mo_only + w1chn_100  + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_chn_m <- feols(chn_dif ~ migr_mo + w1chn_100  + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_chn_f_only <- feols(chn_dif ~ migr_fa_only + w1chn_100  + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_chn_f <- feols(chn_dif ~ migr_fa + w1chn_100  + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl) 
did_adj_unres_chn_both <- feols(chn_dif ~ migr_both + w1chn_100  + age + ch.girl + ch.health + ch.rural + ch.sibling + ch.illness + ch.preschool + ch.skip.grd + ch.repeat.grd +ch.boarding + ch.ethn.minority + par.edu.year + par.ccp + home.finances + home.book + home.internet | schids, data = d_cl)
```
